---
title: "extroversion"
author: "sarah haney"
date: "2025-04-08"
output: html_document
---

```{r setup}
library(dplyr)
library(readr)
library(ggplot2)

extro <- read_csv(
  '../data/TheExtraversion Effectother.csv',
  show_col_types = FALSE)

```


```{r, reversing the scoring for question 33, 36, and 38}

extro <- extro[-3, ]

reverse_score <- function(x) {
  7 - x  # Since max + min = 7 for a 1-6 scale
}

columns_to_convert <- c("Q32", "Q33", "Q34", "Q35", "Q36", "Q37", "Q38", "Q39")
extro[columns_to_convert] <- lapply(extro[columns_to_convert], function(x) as.numeric(as.character(x)))

# Applying reverse scoring
extro <- extro %>%
  mutate(
    Q33 = reverse_score(Q33),
    Q36 = reverse_score(Q36),
    Q38 = reverse_score(Q38)
  )
```

#Hypotheses

Our hypothesis
  H1- We expect a relationship between hours spent socializing with strangers and quality of day
  H2- We expect a relationship between the extroversion scale and the relationship of hours and quality
  
# Descriptive statistics

## Relationship between hours and quality
```{r}


# Assigning hours based on non-NA values across different columns
extro <- extro %>%
  mutate(HoursSocializing = case_when(
    !is.na(Q22_1) ~ 4,
    !is.na(Q24_1) ~ 2,
    !is.na(Q28_1) ~ 0
  ))

# Convert columns to numeric, handling potential non-numeric characters
extro$Q22_1 <- as.numeric(as.character(extro$Q22_1))
extro$Q24_1 <- as.numeric(as.character(extro$Q24_1))
extro$Q28_1 <- as.numeric(as.character(extro$Q28_1))

# Check for errors or warnings after conversion

```

```{r create column QualityOf Day}
# Create a new column 'QualityOfDay' that is the average of non-NA day ratings
extro <- extro %>%
  mutate(QualityOfDay = rowMeans(select(., Q22_1, Q24_1, Q28_1), na.rm = TRUE))

```


```{r}
cor.test(
    ~ QualityOfDay + HoursSocializing,
    data = extro,
    method = 'spearman',
    exact = FALSE
  )
```

## Extroversion Score 
```{r extroversion score}
extro <- extro %>%
  rowwise() %>%
  mutate(
    Extraversion_Score = mean(c(Q32, Q33, Q34, Q35, Q36, Q37, Q38, Q39), na.rm = TRUE)
  )
```

# Correlation of Extroversion and Day

```{r extroversion score- day -correlation for Q22_1}
extro$Q22_1 <- as.numeric(extro$Q22_1)
extro <- extro[!is.na(extro$Q22_1) & extro$Q22_1 != "", ]

cor.test(
    ~ Extraversion_Score + Q22_1,
    data = extro,
    method = 'spearman',
    exact = FALSE
  )
```

```{r extroversion score- day -correlation for Q24_1}
extro$Q24_1 <- as.numeric(extro$Q24_1)
extro <- extro[!is.na(extro$Q24_1) & extro$Q24_1 != "", ]

cor.test(
    ~ Extraversion_Score + QualityOfDay,
    data = extro,
    method = 'spearman',
    exact = FALSE
  )
```

```{r extroversion score- day -correlation for Q28_1}
extro$Q28_1 <- as.numeric(extro$Q28_1)
extro <- extro[!is.na(extro$Q28_1) & extro$Q28_1 != "", ]

cor.test(
    ~ Extraversion_Score + Q28_1,
    data = extro,
    method = 'spearman',
    exact = FALSE
  )
```

```{r linear model}
library(dplyr)

extro %>%
  # Make sure all predictor and response variables are numeric
  mutate(
    QualityOfDay = as.numeric(QualityOfDay),
    HoursSocializing = as.numeric(HoursSocializing),
    Extraversion_Score = as.numeric(Extraversion_Score)  # Correct column name
  ) %>%
  # Fit the linear model
  lm(QualityOfDay ~ HoursSocializing + Extraversion_Score, data = .) %>%
  # Show the results
  summary()

```
QualityOfDay = 124.49+2.93(HoursSocializing)−14.53(ExtraversionScore)

this shows that as hours of socializing increase the score goes up 2.93 and the more extraverted someone is the worse the score is in general



