---
title: "extroversion"
author: "sarah haney"
date: "2025-04-08"
output:
  word_document: default
  html_document: default
---

```{r setup}
library(dplyr)
library(readr)
library(ggplot2)

extro <- read_csv(
  '../data/TheExtraversion Effectother.csv',
  show_col_types = FALSE)

```


```{r, reversing the scoring for question 33, 36, and 38}

extro <- extro[-3, ]

reverse_score <- function(x) {
  7 - x  # Since max + min = 7 for a 1-6 scale
}

columns_to_convert <- c("Q32", "Q33", "Q34", "Q35", "Q36", "Q37", "Q38", "Q39")
extro[columns_to_convert] <- lapply(extro[columns_to_convert], function(x) as.numeric(as.character(x)))

# Applying reverse scoring
extro <- extro %>%
  mutate(
    Q33 = reverse_score(Q33),
    Q36 = reverse_score(Q36),
    Q38 = reverse_score(Q38)
  )
```

#Hypotheses

Our hypothesis
  H1- We expect a relationship between hours spent socializing with strangers and quality of day
  H2- We expect a relationship between the extroversion scale and the relationship of hours and quality
  
# Descriptive statistics

## Relationship between hours and quality
```{r}

# Assigning hours based on non-NA values across different columns
extro <- extro %>%
  mutate(HoursSocializing = case_when(
    !is.na(Q22_1) ~ 4,
    !is.na(Q24_1) ~ 2,
    !is.na(Q28_1) ~ 0
  ))

# Convert columns to numeric, handling potential non-numeric characters
extro$Q22_1 <- as.numeric(as.character(extro$Q22_1))
extro$Q24_1 <- as.numeric(as.character(extro$Q24_1))
extro$Q28_1 <- as.numeric(as.character(extro$Q28_1))

# Check for errors or warnings after conversion

```

```{r create column QualityOf Day}
# Create a new column 'QualityOfDay' that is the average of non-NA day ratings
extro <- extro %>%
  mutate(QualityOfDay = rowMeans(select(., Q22_1, Q24_1, Q28_1), na.rm = TRUE))

```


```{r}
cor.test(
    ~ QualityOfDay + HoursSocializing,
    data = extro,
    method = 'spearman',
    exact = FALSE
  )
```
There is a weak, positive correlation bewteen the quality of day in general and the hours of socializing but this weak correlation is not statistically significant 

## Extroversion Score 
```{r extroversion score}
extro <- extro %>%
  rowwise() %>%
  mutate(
    Extraversion_Score = mean(c(Q32, Q33, Q34, Q35, Q36, Q37, Q38, Q39), na.rm = TRUE)
  )
```

## Count of Extroverts in data
```{r}
sum(extro$Extraversion_Score > 3, na.rm = TRUE)
```

# Correlation of Extroversion and Day

```{r}
run_extro_correlation <- function(data, hour_column) {
  data[[hour_column]] <- as.numeric(data[[hour_column]])
  subset <- data[!is.na(data[[hour_column]]) & data[[hour_column]] != "", ]
  
  cor.test(
    ~ Extraversion_Score + QualityOfDay,
    data = subset,
    method = "spearman",
    exact = FALSE
  )
}
```

```{r}
run_extro_correlation(extro, "Q22_1")
run_extro_correlation(extro, "Q24_1")
run_extro_correlation(extro, "Q28_1")
```

there is a weak, positive correlation between extroversion score and the quality of the day, but it is not statistically significant when the socialising hour is 4 

there is a very weak, negative correlation between extroversion score and quality of day when there is socialization for 2 hours. this weak correlation is not statistically significant 

there is a strong, statistically significant negative correlation between extroversion score and the quality of day score given by participants when the socializing hours are 0  

## Linear Model
```{r linear model}
library(dplyr)

extro %>%
  # Make sure all predictor and response variables are numeric
  mutate(
    QualityOfDay = as.numeric(QualityOfDay),
    HoursSocializing = as.numeric(HoursSocializing),
    Extraversion_Score = as.numeric(Extraversion_Score)  # Correct column name
  ) %>%
  # Fit the linear model
  lm(QualityOfDay ~ HoursSocializing * Extraversion_Score, data = .) %>%
  # Show the results
  summary()

```
QualityOfDay = 104.343 + 3.615(HoursSocializing) − 9.565(ExtraversionScore)

This shows that as hours of socializing increase the score goes up 3.615 and the more extraverted someone is the worse the score is in general

*limitation is the sample because its not representative of any sample, extroverts are way to over represented
*one study says that study shouldn't have an effect but we can see that it might have

## Age as a control
```{r}
model <- lm(QualityOfDay ~ HoursSocializing * Extraversion_Score + Q4, data = extro)
summary(model)

```
Although the model did not reach conventional levels of significance (p = .11), it explained a substantial amount of variance in Quality of Day (R² = .42).
There was a trend suggesting that greater hours socializing were associated with lower Quality of Day, but that this relationship was moderated by Extraversion, such that more extraverted individuals experienced a less negative impact.
participants aged 26-29 and 30-33 also tended to report higher Quality of Day compared to younger participants, though these differences did not reach statistical significance.


## Gender controlled
```{r}
model <- lm(QualityOfDay ~ HoursSocializing * Extraversion_Score + Q41, data = extro)
summary(model)
```
The regression model including hours socializing, extraversion, their interaction, and gender did not significantly predict quality of day, though there was a trend for higher extraversion to be associated with lower quality of day (p = .073). 
--- basically gender doesnt really seem to have an impact on this at all 

## Education Controlled
```{r}
model <- lm(QualityOfDay ~ HoursSocializing * Extraversion_Score + Q6, data = extro)
summary(model)
```
In a model controlling for education level, higher extraversion was significantly associated with lower Quality of Day (p = .018), and there was a significant interaction between hours socializing and extraversion (p = .050), such that greater extraversion reduced the negative effects of socializing hours on Quality of Day.
